### 2016-01-25 letter.tex
begin letter to editor/referee and fix one tiny typo in draft

\documentclass[12pt]{article}

\addtolength{\textwidth}{1in}
\addtolength{\oddsidemargin}{-.5in} %left margin
\addtolength{\evensidemargin}{-.5in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{18.0pt}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\linespread{1.2}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\lhead{Trade Agreements in the Shadow of Lobbying}
%\chead{}
\rhead{Kristy Buzard}
%\fancyfoot[C]{}
%\fancyhead[R]{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}
\usepackage[pdftex]{graphicx}
\setcounter{secnumdepth}{-1}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont}
\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}

\begin{document}

\begin{center}
\large Response to Editor and Referee \\
\normalsize February 2, 2016
\end{center}

Editor's point $\#$7: Thank you for pointing out that the second order condition holds everywhere and that the proof of what was formerly Lemma 1 can be drastically simplified. I moved the discussion of the second order condition into the text and integrated it with the simple result on the increasingness of trade war tariff in the political economy weight. As this did not need to be referenced elsewhere and no longer requires a proof in the appendix, it is no longer labeled Lemma 1. Likewise, Lemma 0 is no longer necessary and has been removed. Note that this advances the numbering of the lemmas in the revised draft relative to the previous draft.

\end{document}