### 2016-01-29 game-tree.tex
add game-tree.tex and start of code for extensive form into draft.tex

\documentclass[10pt]{article}

\addtolength{\textwidth}{1.3in}
\addtolength{\oddsidemargin}{-.65in} %left margin
\addtolength{\evensidemargin}{-.65in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
%\linespread{1.5}
\usepackage{setspace}
\doublespacing

\usepackage{titling}

\usepackage[pdftex]{graphicx}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont,pdfsync}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{tikz}

% disables chapter, section and subsection numbering
%\setcounter{secnumdepth}{-1}

\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
\usepackage[hang,flushmargin]{footmisc}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{assumption}{Assumption}
\newtheorem{fact}{Fact}
\newtheorem{result}{Result}

\newcommand{\ve}{\theta}
\newcommand{\ta}{\theta}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\bte}{\bm{\tau^E}}
\newcommand{\btn}{\bm{\tau^n}}
\newcommand{\ga}{\gamma}

\begin{document}

\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5,fill=white}
%non node/.style={circle,draw,inner sep=.1,fill=black}
}
\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
% The Tree
\node(0)[solid node,label=above:{$E,E^*$}]{} 
child{node(1)[solid node,label=below:{$0$}]{}
}
child{[white] node(2)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
child{[black] node(4)[solid node,label=below:{$0$}]{}
}
child{[white] node(5)[hollow node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
child{[black] node(7)[solid node,label=below:{$0$}]{} 
}
child{[white] node(8)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
child{[black] node(10)[solid node,label=below:{}]{} }
child{[black] node(11)[solid node]{} 
}
child{[white] node(12)[solid node,xshift=12]{}  %note that you need to adjust the yshift if you change the sibling distance
child{[black] node(13)[solid node,label=below:{}]{} }
child{[black] node(14)[solid node,label=above:{$\infty$}]{} }
}
 edge from parent node[black, xshift=23]{$\ta$} %note that you need to adjust the yshift if you change the level distance
}
child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
 edge from parent node[black, xshift=23]{$e_b$} %note that you need to adjust the yshift if you change the level distance
}
child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
 edge from parent node[black, xshift=23]{$\bta$} %note that you need to adjust the yshift if you change the level distance
}
child{node(3)[solid node,label=above:{$\infty$}]{}
}
;
% information set
    \draw[solid,bend right](1)to(3);
		\draw[solid,bend right](4)to(6);
		\draw[solid,bend right](7)to(9);
\end{tikzpicture}

\end{document}

%child{[black] node(4)[solid node,label=below:{$0$}]{} edge from parent node[left]{$C$}}
%child{[black] node(6)[solid node,label=above:{$\infty$}]{} edge from parent node[right]{$D$}}