### 2016-01-31 game-tree.tex
fix game-tree for uncertainty at trade war stage

%\documentclass[10pt]{article}

%\addtolength{\textwidth}{1.3in}
%\addtolength{\oddsidemargin}{-.65in} %left margin
%\addtolength{\evensidemargin}{-.65in}
%\setlength{\textheight}{9in}
%\setlength{\topmargin}{-.5in}
%\setlength{\headheight}{0.0in}
%\setlength{\footskip}{.375in}
%\renewcommand{\baselinestretch}{1.0}
%\usepackage{setspace}
%\doublespacing

%\usepackage{titling}

%\usepackage[pdftex]{graphicx}

%\usepackage[usenames,dvipsnames]{color}
%\usepackage{cite}
%\usepackage{times, verbatim,bm,pifont,pdfsync}
%\usepackage{caption}
%\usepackage{subcaption}

%\usepackage{tikz}

% disables chapter, section and subsection numbering
%\setcounter{secnumdepth}{-1}

%\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}
%\usepackage[hang,flushmargin]{footmisc}

%\usepackage[pdftex,
%bookmarks=true,
%bookmarksnumbered=false,
%pdfview=fitH,
%bookmarksopen=true]{hyperref}

%\newcommand{\ta}{\theta}
%\newcommand{\al}{\alpha}
%\newcommand{\Ta}{\Theta}
%\newcommand{\expect}{\mathbb{E}}
%\newcommand{\Bt}{B(\bm{\tau^a})}
%\newcommand{\bta}{\bm{\tau^a}}
%\newcommand{\bte}{\bm{\tau^E}}
%\newcommand{\btn}{\bm{\tau^n}}
%\newcommand{\ga}{\gamma}

%\begin{document}

\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5,fill=white}
%non node/.style={circle,draw,inner sep=.1,fill=black}
}
\begin{tikzpicture}[scale=1.5,font=\footnotesize, grow=right]
% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=10mm,sibling distance=10mm]
\tikzstyle{level 2}=[level distance=10mm,sibling distance=10mm]
% The Tree
\node(0)[solid node,label=above:{$E,E^*$}]{} 
child{node(1)[solid node,label=below:{$0$}]{}}
child{[white] node(2)[solid node,xshift=12,label=above:{L}]{}  
	child{[black] node(4)[solid node,label=below:{$0$}]{}}
	child{[white] node(5)[hollow node,xshift=12]{} 
		child{[black] node(7)[solid node,label=below:{$-\infty$}]{}}
		child{[white] node(8)[solid node,xshift=12]{}  
			child{[black] node(10)[solid node]{}  }
			child{[black] node(12)[solid node]{}  
				child[sibling distance=5mm]{[black] node(13)[solid node,label=below:{$0$}]{}}
				child{[white] node(15)[solid node,xshift=7]{} 
					child[sibling distance=5mm]{[black] node(19)[solid node,label=below:{$-\infty$}]{}}
					child{[white] node(21)[solid node,xshift=7]{} 
						child[sibling distance=5mm]{[black] node(16)[solid node,label=below:{$0$}]{} }
						child{[white] node(18)[hollow node,xshift=12,label=left:{}]{}  }
						child[sibling distance=5mm]{[black] node(17)[solid node,label=above:{$\infty$}]{} }
						edge from parent node[black, xshift=20]{$\ta_n$} 
						}
					child[sibling distance=5mm]{[black] node(21)[solid node,label=above:{$\infty$}]{} }
					edge from parent node[black, xshift=20]{$e_n$} 
				}
				child[sibling distance=5mm]{[black] node(14)[solid node,label=above:{$\infty$}]{} }
				}
			edge from parent node[black, xshift=23]{$\ta_b$} 
			}
		child{[black] node(9)[solid node,label=above:{$\infty$}]{} }
		edge from parent node[black, xshift=23]{$e_b$} 
		}
	child{[black] node(6)[solid node,label=above:{$\infty$}]{}}
	edge from parent node[black, xshift=23]{$\bta$} 
	}
child{node(3)[solid node,label=above:{$\infty$}]{}
}
;
% information set
    \draw[solid,bend right](1)to(3);
		\draw[solid,bend right](4)to(6);
		\draw[solid,bend right](7)to(9);
		\draw[solid,bend right](13)to(14);
		\draw[solid,bend right](16)to(17);
		\draw[solid,bend right](19)to(21);
% labels
		\node[above,xshift=5,yshift=5]at(2){$L$}; 
		\node[above,xshift=5,yshift=5]at(5){$N$}; 
		\node[above,xshift=8,yshift=5]at(8){$Leg$}; 
		\node[above,xshift=5,yshift=5]at(12){$L$};
		\node[above,xshift=15,yshift=-17]at(21){$Leg$}; 
		\node[above,xshift=7,yshift=5]at(15){$N$}; 
		\node[left, xshift=-2]at(18){$\tau^n$}; 
		\node[below left,yshift=-10]at(12){Break}; 
		\node[below left,yshift=1]at(10){Abide}; 
		\draw[solid](5)circle(.05cm);
		\draw[solid](0)circle(.15cm);
		\node[right]at (10) {$W_E\left(\bta\right),$};
		\node[right,yshift=-10]at (10) {$\pi\left(\tau^a\right)-e_b,$};
		\node[right,yshift=-20]at (10) {$W_{ML}\left(\ga(e_b,\ta_b),\bta\right)$};
		\node[right]at (18) {$W_E\left(\btn\right),$};
		\node[right,yshift=-10]at (18) {$\pi\left(\tau^n\right)-e_b -e_n,$};
		\node[right,yshift=-20]at (18) {$W_{ML}\left(\ga(e_n,\ta_n),\btn\right)$};

\end{tikzpicture}

%\end{document}