### 2016-04-26 letter-round2.tex
explain choice of symmetry in letter; closes #26

\documentclass[12pt]{report}

\addtolength{\textwidth}{1in}
\addtolength{\oddsidemargin}{-.5in} %left margin
\addtolength{\evensidemargin}{-.5in}
\setlength{\parindent}{0pt}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{-.5in}
\setlength{\headheight}{18.0pt}
\setlength{\footskip}{.375in}
\renewcommand{\baselinestretch}{1.0}
\linespread{1.2}
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\lhead{Trade Agreements in the Shadow of Lobbying}
%\chead{}
\rhead{Kristy Buzard}
%\fancyfoot[C]{}
%\fancyhead[R]{}

\usepackage[pdftex,
bookmarks=true,
bookmarksnumbered=false,
pdfview=fitH,
bookmarksopen=true]{hyperref}
\usepackage[pdftex]{graphicx}
\setcounter{secnumdepth}{-1}

\usepackage[usenames,dvipsnames]{color}
\usepackage{cite}
\usepackage{times, verbatim,bm,pifont}
\usepackage{amsbsy,amssymb, amsmath, amsthm, MnSymbol,bbding}

\begin{document}

\begin{center}
\large Response to Editor and Referee \\
\normalsize Month ???, 2016
\end{center}

\vskip.2in
Thank you. \\

Logistics. Points were not numbered, so pasting them here\\

\textit{The main issue is still point 2 of my previous report: the paper is fundamentally quite interesting but poorly executed. What I would like you to do is write a much more focused, crispier version of the paper. The idea is sufficiently straightforward that it does not need 28 pages of exposition. Please reduce the length of the paper by half.}

\vskip.3in
Stylistic advice:

\begin{itemize}
\renewcommand\labelitemi{-}
	\item \textit{Avoid changes of notation as you go along. This is totally unhelpful to the reader. Please pick one and stick to it throughout the paper.}
	\item \textit{Avoid digressions, lengthy discussions, and overreach: you stretch your model far too much. Stick to the main story. The paper still degenerates far too often into a series of discussions that bring little additional insight.}
	\item \textit{Section 2.4 either belongs to section 1.1 or should move as section 1.2. Section 2.4 is far too long.}
	\item \textit{Your recall stuff far too often. Not all readers suffer from memory losses.}
\end{itemize}

\vskip.3in
Some comments on substance, chronologically:
\begin{itemize}
\renewcommand\labelitemi{-}
	\item \textit{Assuming that lobbies represent only import-competing interests is a strong assumption and it requires theoretical and/or empirical justification.}
	\item \textit{Your model is static, so claiming that it rationalizes the four-year delay and eventual ratification of the US-SK FTA is a stretch. In particular, the interactions with automakers suggests a that a bargaining game under incomplete information would be more appropriate.}
  \item \textit{Your assumption that the identity of the median voter is not known a priori is standard in probability voting models, so you may want to stress the analogy between your setting and such models.}
  \item \textit{You claim that trade agreements act as a domestic political commitment device. To make this point formally, it would be nice to establish conditions under which, in equilibrium, “fast track authority” dominates cherry picking ex ante (in a version of your model with several import competing sectors).} \\
		I believe that I have already made this point formally in Result 3. The proof invokes Result 2 and Lemma 3 as it combines both dynamics that are explained intuitively in the text. I have made this more explicit by referring to those results in the text.

		I agree that the idea you propose concerning fast-track authority / TPA is an important one and it is an idea that a colleague and I are exploring in a separate project as I believe it is beyond the scope of the current paper. The extension of the model to one with multiple sectors and lobbies is one that I believe is worthwhile for many purposes. However, the essential idea of the current paper is, as you have judged, `sufficiently straightforward' that in my view the complications involved in the multi-sector model are not warranted.

  \item \textit{I don’t see how the move of Nature (which determines the probabilities o and o* is consistent with the idea that there is uncertainty about who is the median voter (which was part of your motivation). Assuming that o and o* are mutually exclusive is not great. It would be better to assume some timing (rather than have a simultaneous choice). Since the game is symmetric, assuming that Home’s legislature moves before Foreign’s would be without further loss of generality.}
  \item \textit{“the uncertainty about the median legislator’s identity is resolved.” What is this uncertainty really about?}
  \item \textit{The equilibrium selection criterion requires justification and motivation.}
  \item \textit{Assumption 2 does not formalize any intuition: it is just a labelling.} \\
	Quite right. The text has been modified appropriately.
  \item \textit{I feel uneasy about Assumption 3. I would rather have it as an outcome and, if it is not the only possibly one, a discussion of the conditions under which this arises.}
  \item \textit{The full separability of sectors implies very minimal interactions already. Assuming a fully symmetric model is then highly restrictive. How difficult would it be to relax the latter assumption?} \\
	It's not very hard to relax the symmetry of the \textit{equilibrium}, but it requires adding some restrictions so that the impact on the break probability from $\tau^{*a}$ in Lemma 2 does not outweigh that of $\tau^a$ in Lemma 1. I chose symmetry because it's prevalent in the literature (along with separability) and avoids such seemingly ad-hoc assumptions. I also take advantage of the symmetry for the argument establishing existence of the executives' problem and I'm not completely sure the most elegant way with the least additional restrictions to get that without symmetry.

	How hard it would be to make the underlying model asymmetric depends on the particular form of asymmetry. The one such avenue that is regularly followed in this literature is to have political pressure in only one country. That's a simplification so would easily go through. The only other asymmetry I've seen with any regularity is in the size of the economies (one large, one small). I think I might need a restriction on the size of the asymmetry in this case. I'd also have to generalize the bargaining solution in the case of any model asymmetry.

  \item \textit{Expression 6 establishes that the necessary second order condition holds by assumption 1, does not it?} \\
	I agree that this is the case. As you so helpfully pointed this out in your last set of comments, I have included an argument to this effect in the two full sentences following Expression 6. If there is a more direct argument, I would be happy to replace the current one.
  \item \textit{Writing that lobbying expenditures are equal to 0.0007 is meaningless absent units. Are these cents? Trillions of dollars? It would be much more informative if these were expressed as a fraction of equilibrium profits.}
  \item \textit{The last paragraph of section 4 is quite a stretch; is there any evidence in the political science literature that the separation of power is correlated with trade openness among democracies?} \\
	The issue is quite a nuanced, in particular among democracies. I have discussed the empirical evidence on democracies---separate from the claim that I made about moving from autocracy to democracy---from Henisz and Mansfield (2006) in a new paragraph. I have also added a reference to Frye and Mansfield (2003) to support the original claim. I don't consider any of this essential and would be happy to remove either or both paragraphs.
\end{itemize}

Best wishes,\\
\\
\\
Kristy Buzard

\end{document}